Brief Announcement: Tight Bounds for Online Balanced Re-Partitioning

### 2020-02-18 paper/podc20ba.tex
moving PPL alg to the Appendix
feedback

\noindent \textbf{Contributions.}
Our main technical contribution 
is an improved lower bound 
of $\Omega(k\cdot\ell)$ on the competitive ratio of any online deterministic online algorithm 
even in the learning model;
the best known lower bound so far was $\Omega(k)$,
even for a more general model~\cite{repartition-disc}.
We also present an asymptotically optimal, 
$O(k\cdot \ell)$-competitive algorithm
for the restricted model.
For the online model, we present  
an asymptotically optimal,
$\Theta(\ell)$ algorithm for $k=3$;
the best known upper bound 
so far was $O(\ell^2)$~\cite{repartition-disc}.
%
Our algorithms can be distributed
similarly to the approach in~\cite{sigmetrics19_partitioning}.

@@

\noindent \textbf{Preliminaries.}
In this paper, we use the technique of maintaining \emph{(connected) components} of nodes, similar to previous approaches~\cite{repartition-disc}. A~\emph{component} is a subset of frequently communicating nodes.
Initially,
every node is trivially a \emph{singleton} component.
An~algorithm is \emph{component respecting}
if it always keeps nodes that  belong to the same component collocated.
That is,
if the algorithm needs to move a node,
it moves the whole component containing the node.
%We say that the request $(u,v)$ is \emph{external}, if the algorithm is in the configuration, where $u$ and $v$ are located at different clusters.
%This is achieved by opting a sequence of  partitions (a.k.a.~configurations)
%such that frequently communicating nodes (from different clusters)
%are relocated to the same cluster before they inflict too much communication cost.
%A \emph{reconfiguration} (i.e.,\emph{ re-partitioning})
%is performed by migrating nodes between clusters,
%at the total cost of individual node migrations.
%The objective is to jointly minimize communication and migration costs while serving the sequence.

@@

\noindent \textbf{Lower bound.}
The adversary constructs components of nodes, and issues only requests among nodes in the same component.
If an online algorithm splits the components at any moment, the adversary issues requests to split pairs of nodes.
We utilize the fact that the online algorithm does not know the components of the adversary, and it may split some components while evicting nodes.
On the other hand, $\OPT$ never splits the components.

>>>>

\noindent \textbf{Contributions.}
We provide an improved lower bound 
of $\Omega(k\cdot\ell)$ on the competitive ratio of any online deterministic online algorithm 
even in the learning model;
the best known lower bound so far was $\Omega(k)$,
even for a more general model~\cite{repartition-disc}.
We also present an asymptotically optimal, 
$O(k\cdot \ell)$-competitive algorithm
for the restricted model.
For the online model, we present  
an asymptotically optimal,
$\Theta(\ell)$ algorithm for $k=3$;
the best known upper bound 
so far was $O(\ell^2)$~\cite{repartition-disc}.
%
Our algorithms can be distributed
similarly to the approach in~\cite{sigmetrics19_partitioning}.

@@

\noindent \textbf{Preliminaries.}
In this paper, we use the technique of maintaining \emph{(connected) components} of nodes, similar to previous approaches~\cite{repartition-disc}. A~\emph{component} is a subset of frequently communicating nodes.
We use this concept in both algorithms in this paper, but also in the lower bound, to incur high cost for any online algorithm that splits them.
An~algorithm is \emph{component respecting}
if it always keeps nodes that  belong to the same component collocated.
That is,
if the algorithm needs to move a node,
it moves the whole component containing the node.
%We say that the request $(u,v)$ is \emph{external}, if the algorithm is in the configuration, where $u$ and $v$ are located at different clusters.
%This is achieved by opting a sequence of  partitions (a.k.a.~configurations)
%such that frequently communicating nodes (from different clusters)
%are relocated to the same cluster before they inflict too much communication cost.
%A \emph{reconfiguration} (i.e.,\emph{ re-partitioning})
%is performed by migrating nodes between clusters,
%at the total cost of individual node migrations.
%The objective is to jointly minimize communication and migration costs while serving the sequence.

@@

\noindent \textbf{Lower bound.}
The adversary constructs components of nodes, and issues only requests among nodes in the same component.
If an online algorithm splits the components at any moment, the adversary issues requests to split pairs of nodes.
We utilize the fact that the online algorithm does not know the components of the adversary, and it may split some components while evicting nodes.
On the other hand, $\OPT$ initially collocates all components and never splits them again.
